Detection of DCIS and IDC in Whole-Slide\\ H\&E Stained Breast Histopathology Images Using Stacked Convolutional Neural Networks

### 2016-10-07 tex/article.tex
Gen 1024, 512 graphs and configs for these

@@

@@

\begin{figure*}[!t]
\centering
\hspace{-0.1in}\includegraphics[width=6.5in]{system_overview}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
\caption{Overview of the proposed system for labeling whole-slide images. (a) Original WSI of breast tissue. (b) Resulting probability map of applying the CNN in a patch-based fashion. (c) Output of the system; a single label for the whole-slide image (either benign, DCIS or IDC).}
\end{figure*}

@@

@@

@@

\subsubsection{Wide Residual Networks}
Residual network was the winner of the 2015 ImageNet challenge \cite{resnet}. Other than previous networks that entered this competition, it features no fully connected layer at the end of the network. It was the deepest succesful architecture to date featuring up to 152 layers. It can get away with this depth by using residual learning blocks, where the layers learn residual functions with relation to the input instead of learning unreferenced functions. To the extreme, it makes it easier to learn the identity function if that is optimal for that layer.

@@

The architecture is a recipe that has three main design choices, the $N$ value that determines the depth, the $k$ value that determines the width (the amount of filters) and the type of ResNet block. Here, we use $N=4$, $k=2$, which was empirically evaluated to fit a decent batch size and train fast. The ResNet block type is detailed in figure \ref{fig_resnet_block}, TODO more about this block. The resulting architecture is shown in figure \ref{fig_resnet}.

@@

\begin{figure*}[!t]
\centering
\subfloat[{Wide Residual Network architecture. \newline \emph{RB} stands for ResNet Block.}]{\includegraphics[width=2.20in]{architecture_wide_resnet}%
\label{fig_resnet}}
\subfloat[{Architecture of the resnet blocks. \newline The $\otimes$ node indicates an elementwise sum.}]{\hspace{0.8cm}\includegraphics[width=1.8in]{architecture_resnet_block}%
\label{fig_resnet_block}}
\subfloat[VGG-16 model architecture. \newline Dropout with probability of 0.5 was used in the fully connected layers.]{\hspace{0.98cm}\includegraphics[width=2.20in]{architecture_vgg}%
\label{fig_vgg}}
\caption{Architectures used for 224x224 patch classification. The 1x1 convolution layer in the resnet blocks is only present in the blocks where the input is downsampled, which is every first block with a higher amount of filters in the Wide ResNet architecture (figure \ref{fig_resnet}).}
\label{fig_architectures}
\end{figure*}

@@

\begin{figure}[!t]
\centering{}
\includegraphics[width=2.5in]{architecture_stacked}
\caption{Architecture of the stacked network. The weights of the components with the dotted outlines are taken from the previously trained 224x224 patch model, and are no longer updated (these are frozen).}
\label{fig_stacked}
\end{figure}

@@

\section{Whole-slide image labeling}

@@

\begin{figure}[!t]
\centering{}
\hspace{-0.3cm}\includegraphics[width=3.5in]{2class_performance}
\caption{Performance of networks trained on the two class problem of benign versus cancerous patches. The dashed line shows the performance on the images from the train set it was shown that epoch.}
\label{fig_2class}
\end{figure}

\begin{figure}[!t]
\centering{}
\hspace{-0.3cm}\includegraphics[width=3.5in]{3class_performance}
\caption{Performance of networks trained on all three classes. The dashed line shows the performance on the images from the train set it was shown that epoch. The validation accuracy of the wide resnet is especially noisy.}
\label{fig_3class}
\end{figure}

@@

\begin{table}[!t]
%% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.1}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Best epoch patch-level accuracy of 768x768 stacked networks}
\label{table_results_stacked}
\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
\begin{tabular}{|llr|}
\hline
\textsc{Labels}&\textsc{Stacked on}&\textsc{Accuracy}\\
\hline
\textit{Benign, Cancer}&&\\
&2 Class WRN-4-2&  123\\
&3 Class WRN-4-2& 123\\
\hline
\textit{Benign, DCIS, IDC}&&\\
&2 Class WRN-4-2&  123\\
&3 Class WRN-4-2& 123\\
\hline
\end{tabular}
\end{table}

>>>>

\usepackage{pbox}

@@

\subsection{DCIS and IDC}
A lesion is considered to be a DCIS when there is a proliferation of malignant epithelial cells within the breast parenchymal structures. When these cells are no longer contained within this basement membrane, the carcinoma is invasive and is classified as an invasive ductal carcinoma (IDC) \cite{ellis2010intraductal}.

\begin{figure*}[!t]
\centering
\includegraphics[width=7in]{lesion_examples}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
\caption{Examples from the three classes we aim to distinguish. (a) Normal tissue and benign lesions. From top-to-bottom, left-to-right: lymphocytes?? with connective tissue surrounding it, mostly fatty tissue, healthy mammal ducts, a ductal hyperplasia which is very similar in appearance to a DCIS lesion. (b) Ductal carcinoma in situ (DCIS), note that the tumor cells are contained within the duct structure. (c) Invasive ductal carcinoma (IDC).}
\end{figure*}

@@

\begin{figure*}[!t]
\centering
\hspace{-0.08in}\includegraphics[width=6in]{system_overview}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
%\label{fig_second_case}}
\vspace{-0.1cm}\caption{Overview of the proposed system for labeling whole-slide images. (a) Original WSI of breast tissue. (b) Resulting probability map of applying the CNN in a patch-based fashion. (c) Output of the system; a single label for the whole-slide image (either benign, DCIS or IDC).}
\end{figure*}

@@

	% Section dense prediction

@@

\begin{table}[!t]
%% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.05}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Distribution of dataset consisting of 235 whole-slide images}
\label{table_dataset}
\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
\begin{tabular}{|llccc|}
\hline
\textsc{Label}&\textsc{Lesion type}&\hspace{-0.4cm}\textsc{WSI Count}&&\\
\hline
&&\textit{Train}&\hspace{-0.24cm}\textit{Validation}&\textit{Test}\\
\textit{Benign}&&&&\\
&None (normal)& ?&&\\
&Benign lesions& ? &&\\
%&Ductal hyperplasia& ?&&\\
%&Adenosis& ?&&\\
%&Fibrosis& ?&&\\
%&Duct ectasia& ?&&\\
%&Fibrocystic ectasia& ?&&\\
%&Hamartoma& ?&&\\
%\rule{0pt}{3ex}
%&\pbox{2.8cm}{Pseudoangiomatous \\ \-\ \-\ \-\ stromal hyperplasia}& ?&&\\
%\rule{0pt}{3ex}
%&\pbox{2.8cm}{Sclerosing  \\ \-\ \-\ \-\ lobular hyperplasia}& ?&&\\
%\rule{0pt}{2ex}
%&Fibroadenoma& ?&&\\
%&Mixed benign lesions& ?&&\\
&\textit{Total}& 54&20&34\\

%\hline
\textit{DCIS}&&&&\\
&Grade I DCIS&  ?&&\\
&Grade II DCIS& ?&&\\
&Grade III DCIS& ?&&\\
&\textit{Total}& 42&12&20\\
%\hline
\textit{IDC}&&&&\\
&IDC& ?&&\\
&\textit{Total}& 30&9&14\\

\textit{\textbf{Total}}&&&&\\
&&\textbf{126}&\textbf{41}&\textbf{68}\\
\hline
\end{tabular}
\end{table}

@@

\subsubsection{Wide Residual Network}
Residual network was the winner of the 2015 ImageNet challenge \cite{resnet}. Other than previous networks that entered this competition, it features no fully connected layer at the end of the network. It was the deepest succesful architecture to date featuring up to 152 layers. It can get away with this depth by using residual learning blocks, where the layers learn residual functions with relation to the input instead of learning unreferenced functions. To the extreme, it makes it easier to learn the identity function if that is optimal for that layer.

@@

The architecture is a recipe that has three main design choices, the $N$ value that determines the depth, the $k$ value that determines the width (the amount of filters) and the type of ResNet block. Here, we use $N=4$, $k=2$, which was empirically evaluated to fit a decent batch size and train fast. The ResNet block type is detailed in figure \ref{fig_resnet_block}, this is projection block type B as detailed in the original paper. The resulting architecture is shown in figure \ref{fig_resnet}.

@@

\begin{figure*}[!t]
\centering
	\subfloat[{Wide Residual Network architecture. \newline \emph{RB} stands for ResNet Block.}]{\includegraphics[width=2.20in]{architecture_wide_resnet}%
\label{fig_resnet}}
\subfloat[{Architecture of the resnet blocks. \newline The $\otimes$ node indicates an elementwise sum.}]{\hspace{0.8cm}\includegraphics[width=1.8in]{architecture_resnet_block}%
\label{fig_resnet_block}}
\subfloat[VGG-16 model architecture. \newline Dropout with probability of 0.5 was used in the fully connected layers.]{\hspace{0.98cm}\includegraphics[width=2.20in]{architecture_vgg}%
\label{fig_vgg}}
\vspace{-0.12cm}
\caption{Architectures used for 224x224 patch classification. The 1x1 convolution layer in the resnet blocks is only present in the blocks where the input is downsampled, which is every first block with a higher amount of filters in the Wide ResNet architecture (figure \ref{fig_resnet}).}
\label{fig_architectures}
\end{figure*}

@@

\begin{figure}[!t]
\centering{}
\includegraphics[width=2.2in]{architecture_stacked}
\vspace{-0.15cm}\caption{Architecture of the stacked network. The weights of the components with the dotted outlines are taken from the previously trained 224x224 patch model, and are no longer updated (these are frozen).}
\label{fig_stacked}
\end{figure}

@@

\section{Whole-slide image labeling}
A secondary goal is providing a label for the whole-slide image. We create a dense prediction using the stacked convolutional neural network trained previously and extract features from this probability map. We then use these features to train a classifier for classifying this label for whole-slide image.

We will treat it as a multiclass problem; we will assign one label to a whole-slide image. In actuality this problem is multilabel. A slide with invasive (IDC) lesions, may also contain DCIS lesions. Also, of course, these slides will also likely contain benign lesions. If a slide contains any IDC, it should be labeled IDC. If it contains DCIS but not IDC, it should be labeled DCIS. Otherwise, the label is benign.

We tweak and evaluate the models by applying 10-fold cross validation on the combined set of train and validation whole slide images. As a final evaluation, we classify images from the held-out test subset of the data. This data has not been used to train any of the models in the whole pipeline, and will thus be a good measure of the generalisability of the method. Refer to table \ref{table_dataset} for the amount of whole-slide images per class.

The features we extract are the following:

@@

\begin{figure}[!t]
\centering{}
\hspace{-0.3cm}\includegraphics[width=3.5in]{2class_performance}
\vspace{-0.35cm}\caption{Performance of networks trained on the two class problem of benign versus cancerous patches. The dashed line shows the performance on the images from the train set it was shown that epoch.}
\label{fig_2class}
\end{figure}

\begin{figure}[!t]
\centering{}
\hspace{-0.3cm}\includegraphics[width=3.5in]{3class_performance}
\vspace{-0.35cm}\caption{Performance of networks trained on all three classes. The dashed line shows the performance on the images from the train set it was shown that epoch. The validation accuracy of the wide resnet is especially noisy.}
\label{fig_3class}
\end{figure}

\begin{figure}[!t]
\centering{}
\hspace{-0.3cm}\includegraphics[width=3.5in]{768_performance}
\vspace{-0.35cm}\caption{Performance of stacked networks, either on the 2 class WRN-4-2 and the 3 class WRN-4-2 model.}
\label{fig_stack}
\end{figure}

\begin{figure}[!t]
\centering{}
\hspace{-0.3cm}\includegraphics[width=3.5in]{patch_sizes_performance}
\vspace{-0.35cm}\caption{Performance of networks with different input sizes, all stacked on the 3 class WRN-4-2 model.}
\label{fig_stack_different_input_sizes}
\end{figure}

\begin{table}[!t]
%% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.1}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Regions of interest (ROI) per class}
\label{table_roi_count}
\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
\begin{tabular}{|lcc|}
\hline
\textsc{Label}&\textsc{ROI Count}&\\
\hline
&\textit{Train}&\textit{Validation}\\
Benign&5097& 795\\
%\hline
DCIS&1092&  151\\
%\hline
IDC&2307& 432\\
\hline

\end{tabular}
\end{table}

@@

\begin{table}[!t]
%% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.1}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
\caption{Best epoch patch-level accuracy of stacked networks}
\label{table_results_stacked}
\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
\begin{tabular}{|llr|}
\hline
\textsc{Input patch size}\pbox{2cm}{}&\textsc{Stacked on}&\textsc{Accuracy}\\
\hline
\textit{512x512}&&\\
&3 Class WRN-4-2& 123\\
%\hline
\textit{768x768}&&\\
&2 Class WRN-4-2&  123\\
&3 Class WRN-4-2& 123\\
%\hline
\textit{1024x1024}&&\\
&3 Class WRN-4-2& 123\\
\hline

\end{tabular}
\end{table}
